---
title: "Lab 2 Mireia"
author: "Mireia Valle"
date: "16/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Install packages

#install.packages ("naniar")
#install.packages ("VIM")
#install.packages ("skimr")
#install.packages ("ggfortify")
#install.packages ("shinythemes")

#load libraries

library (tidyverse)
library (janitor)
library(naniar)
library(VIM)
library(skimr)
library(ggfortify)
library (here)
```

##1. Get some data
```{r}
ca_pb <- read_csv(here("data", "ca_pollution_burden.csv"))

ca_dem <- read_csv (here("data", "ca_census_demographics_2010.csv"))
```



##2. Some data cleaning
```{r}
ca_pb_nopct <- ca_pb %>% 
  clean_names() %>% 
  select(-contains ("perc")) %>% 
  select(-contains ("pctl")) %>% 
  select(-latitude, -longitude)

ca_dem_clean <- ca_dem %>% 
  clean_names()
  
```
##3. Pollution burden subset
```{r}

ca_pb_subset <- ca_pb_nopct %>% 
  select (ozone, solid_waste, asthma:housing_burden)
```

  
##4. PCA
```{r}
#pb_pca <- prcomp (ca_pb_subset, scale=TRUE)
#it gives an error because the data has NAs
```



#Aside: exploring missining values
```{r}
summary(ca_pb_subset)

gg_miss_var(ca_pb_subset)

matrixplot(ca_pb_subset)
```


#Remove NA (complete deletion)
```{r}
ca_pb_nona <- ca_pb_subset %>% 
  drop_na()

summary (ca_pb_nona)
```


#Now that we do not have NAs, we can do PCA  
```{r}
my_ca_pca <- prcomp(ca_pb_nona)

my_ca_pca

summary(my_ca_pca)

biplot(my_ca_pca)

my_biplot <- autoplot(my_ca_pca, 
                      colour = NA,
                      loadings.label = TRUE,
                      loadings.label.size = 3,
                      loadings.label.colour = "black",
                      loadings.label.repel = TRUE) +
  theme_minimal()
  

my_biplot

```
##4. PCA for pollution burden + demographics

Join together the ca_dem_clean with ca_pb_nopct

```{r}

ca_df <- ca_dem_clean %>% 
  inner_join(ca_pb_nopct, by= c("census_tract_number" = "census_tract"))

#Now drop NAs by complete case deletion:

ca_nona <- ca_df %>% 
  drop_na()
```

Make a subset of ca_nona that only includes: 
-white_percent
-elderly_65_percent
-pm2_5
-pesticides
-traffic
-asthma
-cardiovascular_disease
-poverty

```{r}

my_sub <- ca_nona %>% 
  select (white_percent, elderly_65_percent, pm2_5, pesticides, traffic, asthma, cardiovascular_disease, poverty)

#PCA:

my_dem_PCA <- prcomp (my_sub, scale =TRUE)

my_dem_PCA

summary (my_dem_PCA)

my_dem_PCA <- autoplot(my_dem_PCA, 
                      colour = NA,
                      loadings.label = TRUE,
                      loadings.label.size = 3,
                      loadings.label.colour = "black",
                      loadings.label.repel = TRUE) +
  theme_minimal()
  
my_dem_PCA 

```



